Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-08-10 Manuscript/manuscriptPS.tex
Add figures delivered by Pavel Buslaev: http://nmrlipids.blogspot.com/2018/07/nmrlipids-iv-first-draft-of-manuscript.html?showComment=1533838323396#c5055037960206662973

\author{Tiago M. Ferreira}
\affiliation{Halle, Germany}
\author{Ivan Gushchin}
\affiliation{Moscow Institute of Physics and Technology}
\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{Batuhan Kav}
\affiliation{Potsdam, Germany}
\author{Jesper J. Madsen}
\affiliation{Department of Chemistry, The University of Chicago, Chicago, Illinois 60637, United States of America}
\author{Markus Miettinen}
\affiliation{Potsdam, Germany}
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}
\author{Thomas Piggot   \todo{Authorlist is not yet complete}}
\affiliation{Southampton, United Kingdom}
%\author{NMRlipids collaboration}
%\affiliation{nmrlipids.blogspot.fi}

@@

None of the tested models reproduce the experimental headgroup and 
glycerol backbone order parameters of DOPS and POPS within the
experimental error bars (Fig. \ref{HGorderParametersPS}). 
The tested models perform generally less well than the models tested
for PC headgroup in the previous study (Fig. 2 in Ref.~\cite{botan15}),
which is also evindent from the comparison between subjective rankings
of the model quality for the headgroup and glycerol backbone 
(Fig. \ref{comparisonTablePS} and Fig. 4 in Ref.~\cite{botan15}).
Therefore, the models cannot be straightforwardly used to interpret the
structural differences between PC and PS headgroups. 
However, the differences are partially reproduced by the two best performing 
models for the $\alpha$ and $\beta$-carbons of PS headgroups, Slipids and CHARMM36.
Both reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order 
in parameter in the PS headgroups (Fig. \ref{HGorderParametersPS} and Fig. 2 in Ref.~\citenum{botan15}).
Interestingly, the dihedral angle distributions 
%of CHARMM36 and Slipids in 
in these two models share significant similarities in the headgroup region (Fig.~\ref{dihedralsHG}). 
\todo{Notation of dihedrals in Fig.~\ref{dihedralsHG} should be made somehow combatible with the
chemical structures in Fig.~\ref{HGorderParameters} and the discussion should be then finished.}
However, the glycerol backbone order parameters in Slipids model
significantly differ from CHARMM36 results and experiments (Fig. \ref{HGorderParametersPS}),
which can be related to the differences in the dihedral angle distributions 
of C1-C2-C3-O31 and C2-C3-O31-C31 (Fig.~\ref{dihedralsGLY}). 
Similar difference was previously observed for PC lipids \cite{botan15}
and the conformational differences are illustrated in Figure \ref{glycerol_buslaev}.

>>>>

\author{Pavel Buslaev}
\affiliation{Moscow Institute of Physics and Technology}
\author{Tiago M. Ferreira}
\affiliation{Halle, Germany}
\author{Ivan Gushchin}
\affiliation{Moscow Institute of Physics and Technology}
\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{Batuhan Kav}
\affiliation{Potsdam, Germany}
\author{Jesper J. Madsen}
\affiliation{Department of Chemistry, The University of Chicago, Chicago, Illinois 60637, United States of America}
\author{Markus Miettinen}
\affiliation{Potsdam, Germany}
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}
\author{Thomas Piggot   \todo{Authorlist is not yet complete}}
\affiliation{Southampton, United Kingdom}
%\author{NMRlipids collaboration}
%\affiliation{nmrlipids.blogspot.fi}

@@

\todo{New figures are delivered by Pavel Buslaev. This should be rewritten.}
None of the tested models reproduce the experimental headgroup and 
glycerol backbone order parameters of DOPS and POPS within the
experimental error bars (Fig. \ref{HGorderParametersPS}). 
The tested models perform generally less well than the models tested
for PC headgroup in the previous study (Fig. 2 in Ref.~\cite{botan15}),
which is also evindent from the comparison between subjective rankings
of the model quality for the headgroup and glycerol backbone 
(Fig. \ref{comparisonTablePS} and Fig. 4 in Ref.~\cite{botan15}).
Therefore, the models cannot be straightforwardly used to interpret the
structural differences between PC and PS headgroups. 
However, the differences are partially reproduced by the two best performing 
models for the $\alpha$ and $\beta$-carbons of PS headgroups, Slipids and CHARMM36.
Both reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order 
in parameter in the PS headgroups (Fig. \ref{HGorderParametersPS} and Fig. 2 in Ref.~\citenum{botan15}).
Interestingly, the dihedral angle distributions 
%of CHARMM36 and Slipids in 
in these two models share significant similarities in the headgroup region (Fig.~\ref{dihedralsHG}). 
However, the glycerol backbone order parameters in Slipids model
significantly differ from CHARMM36 results and experiments (Fig. \ref{HGorderParametersPS}),
which can be related to the differences in the dihedral angle distributions 
of C1-C2-C3-O31 and C2-C3-O31-C31 (Fig.~\ref{dihedralsGLY}). 
Similar difference was previously observed for PC lipids \cite{botan15}
and the conformational differences are illustrated in Figure \ref{glycerol_buslaev}. 
\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/figS8.png}
  \caption{\label{HGstructuresPS}
    Overlayed snapshots from glycerol backbone and headgroup region from 
    different simulations of PS lipids.
  }
\end{figure*}

### 2018-08-10 Manuscript/manuscriptPSsuppl.tex
Add figures delivered by Pavel Buslaev: http://nmrlipids.blogspot.com/2018/07/nmrlipids-iv-first-draft-of-manuscript.html?showComment=1533838323396#c5055037960206662973

\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}

\author{NMRlipids collaboration \todo{Authorship query to be sent soon.}}
\affiliation{nmrlipids.blogspot.fi}

@@

\todo{This is to be removed once the new data is available:
https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/16}

@@

\pagebreak
\section{Dihedrals}
\begin{figure*}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/diheds_pops1.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops2.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops3.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops4.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops5.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops6.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops7.png}
  \caption{\label{dihedralsGLY}
    Dihedral angle distributions of bonds from phosphate to acyl chain carbonyls from different simulation models.
  }
\end{figure*}

\begin{figure*}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/diheds_pops8.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops9.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops10.png}
  \includegraphics[width=8.0cm]{../Figs/diheds_pops11.png}
  \caption{\label{dihedralsHG}
    Dihedral angle distributions of bonds from phosphate to headgroup from different simulation models.
  }
\end{figure*}

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/glycerol_buslaev.png}
  \caption{\label{glycerol_buslaev}
    Snapshots overlayed from different simulations for glycerol backbone region
    by Pavel Buslaev.
  }
\end{figure}
\pagebreak

>>>>

\author{Pavel Buslaev}
\affiliation{Moscow Institute of Physics and Technology}
\author{Tiago M. Ferreira}
\affiliation{Halle, Germany}
\author{Ivan Gushchin}
\affiliation{Moscow Institute of Physics and Technology}
\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{Batuhan Kav}
\affiliation{Potsdam, Germany}
\author{Jesper J. Madsen}
\affiliation{Department of Chemistry, The University of Chicago, Chicago, Illinois 60637, United States of America}
\author{Markus Miettinen}
\affiliation{Potsdam, Germany}
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}
\author{Thomas Piggot   \todo{Authorlist is not yet complete}}
\affiliation{Southampton, United Kingdom}

@@

\todo{This is to be removed once the new data is available: \\
https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/16}

@@

\pagebreak
\section{Dihedral angle distributions of the headgroup and glycerol backbone
  regions of PS lipids from different simulation models}

\begin{figure*}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/PS_labels.png}
  \includegraphics[width=16.0cm]{../Figs/figS7.png}
  \caption{\label{dihedralsHG}
    Dihedral angle distributions of headgroup region of PS lipids from different simulation models.
  }
\end{figure*}

\begin{figure*}[]
  \centering
  \includegraphics[width=15.0cm]{../Figs/figS6.png}
  \caption{\label{dihedralsGLY}
    Dihedral angle distributions of glycerol backbone region of PS lipids from different simulation models.
  }
\end{figure*}

\pagebreak
